ShrinkNets

### 2018-01-05 paper/paper.tex
pass up to 3.1 included

\author{Guillaume Leclerc}
\authornote{Visiting Student}
\MITAffiliation
\email{leclerc@mit.edu}

\author{Raul Castro Fernandez}
\MITAffiliation
\email{raulcf@csail.mit.edu}

\author{Samuel Madden}
\MITAffiliation
\email{madden@csail.mit.edu}

@@

When designing Neural Networks, finding the appropriate size (width and depth)
is key. In particular, these hyper parameters have a strong correlation with
over/underfitting \srm{give reference} \gl{I did some litterature review and
  actually this statement is quite controversial, should we rephrase ?}. We
have no reliable way to find them but decades of experimentation led to some
heuristics ~\cite{Bengio2012a} that try to prune that immense space of possible
network sizes and suggest range of values to explore. To select the best
candidates , researchers have used strategies such as random search \gl{Should
  we also move the references here if we remove the pargraph from the related
  work ?}, meta-gradient descent ~\cite{Pedregosa2016} and Parzen Estimators
~\cite{Bergstra2011a} to reduce find good parameters without exhaustively
exploring the entire hyper-parameter space. Although these strategies help with
finding a good set of hyper-parameters, they still require a compute-intensive
search of the space.

@@

In this paper we present a method to automatically find an appropriate network
size from a single parameter, drastically reducing the hyper-parameter
dimensionality. The key idea is to \emph{learn} the right network size at the
same time that the network is learning the main task. For example, for an image
classification task, with our approach we can provide the training data to a
network---without sizing it a priori---and expect to end up with a network that
has learnt the task without overfitting \gl{too strong, we do not really have
  any guarantee on the generalization, how about changing it to: to end up with
  a network that learnt a tradeoff between size and accuracy}. This approach
has two main benefits. First, we do no longer need to choose a network size
before training.  Second, the final network size will be appropriate for the
task at hand, and not larger \gl{Should we talk about the fact that it automatically gets rid of dead neurons ?}. This is important because oversized networks have
a lower inference throughput and higher  memory footprint.

@@

\textbf{Filter Layers: } They have weights in the range $[0,+\infty]$ and are placed after
linear and convolutional layers of traditional deep networks \gl{This is how I
  used it but with the new implementation we can be more creative and put them
  anywhere, should we say "usually placed" insdead ?}. The \textit{Filter
  Layer} takes an input of size $\left(B \times C \times D_1 \times \dots
  \times D_n\right)$, where $B$ is the batch size, $C$ the number of features
(or channels), and $D$ any additional dimension. This structure makes it
compatible with fully connected layers with $n=0$ or convolutional layers with
$n=2$. Their crucial property is a parameter $\theta \in \mathbb{R}^C$, defined
as follows:

@@

Where $\circ$ is the Hadamard product (pointwise multiplication \gl{Should we
  keep only one of them to be more concise ?}), and $\theta$ is expanded in all
dimensions except the second one to match the input size. It is easy to see
that if for any $k$, if $\theta_k \leq 0$, the $k^{\text{th}}$ feature/channel
will forever be $0$. When this happens, we say the Filter layer disables the
neuron. These disabled neurons/channels can be removed from the network without
changing its output (we describe how we perform this removal below). We explain
next how the weights of the Filter Layer are adjusted during training.

@@

%\subsection{The training procedure}
\textbf{Training Procedure: } Once Filter layers are placed in a deep network, 
%To train networks we need start with a substantially oversized network, then we
%insert \textit{Filter Layers}  (usually after every linear or convolutional
%layer except the last one) and we sample their weight from the
%$\text{Uniform}(0, 1)$ distribution. 
we could train it directly and it
would be equivalent to a normal neural network. However, our goal is to find the
smallest network with reasonable performance. We achieve that by introducing
sparsity in the parameters of the \textit{Filter Layers}. Indeed, having a negative component in the $\theta$ parameter of the filter layer permamently disable its associated feature \gl{Maybe redundant ? we talked about that in the previous paragraph}
. To obtain this sparsity, we simply redefine the loss function:

@@

This choice come from the Lasso loss ~\cite{Tibshirani1996} and we use it for the same reason:
it introduces sparsity. The bigger the $\lambda$ the more entries in $\theta$
are set to zer0, and since zero entries in $\theta$ correspond to dead neurons,
lambda effectively control the number of neurons/channels in the entire
network.
Next, we explain how to implement ShrinkNets efficiently.

\textbf{Software architecture\footnote{The code is available as Python/PyTorch
    library on \url{http://github.com/mitdbg/fastdeepnets} \gl{Should we rename
      the repository ?}}: }\textit{Filter Vectors} can easily be implemented in
a few lines of code in many existing deep learning frameworks. However, in
ShrinkNets, we assume that we start with obviously oversized networks. If
disabled neurons are not quickly removed, the overhead might cause the training
process to be significantly slower than classic neural networks. This is why we
want to support what we call garbage collection: removing the weights in the
layers that are responsible (or using) features that are disabled.  \par To be
able to catpure this relationship between layers, we agumented \textit{Pytorch}
~\cite{paszke2017automatic} (the framework we used as the
foundation of our implementation) with a graph structure very similar to the
one available in Keras ~\cite{chollet2015keras}.  In this graph, edges are
effectively event hubs responsible for propagating \textit{feature removal}
events to its endpoints. ShrinkNet layers are designed to emit and react to
these events but also propagate the event further if has an impact at the other
endpoint of the layer (input or output) \gl{What I mean here is: if the event
  comes from the output, it might be propagated to the inputs, and the other
  way around}.  \par This event-based implementation coordinated by edges makes
it very easy to integrate new layers in the library.  We even provide automatic
wrapping for layers from \textit{PyTorch} if they have no internal state. For
more complex ones, we provide utilitaries that
makes the implementation very concise.

\section{Evaluation}
\subsection{Convergence}

To demonstrate that the approach is viable, we will first show that Shrinking
Networks converge properly. For this experiment we trained a one hidden layer
neural network with one filter layer to control the number of hidden units. We
initialized the models with $10000$ neurons and trained them on \texttt{MNIST}
~\cite{Lecun1998} using different regularization factors ($\lambda$). We
summarized the results in \autoref{convergence_plot}. We can see on this plot
is that for a given $\lambda$ the number of hidden units seems to always
converge to some value. It has two implications: it seems that $\lambda$ is, as
we would think, a proxy of the network size (bigger $\lambda$ imply smaller
networks). Secondly, it indicates that the spikes we see in the regular spikes
we see in the loss that occur when neurons are disabled will eventually
disapear because the number of neurons reach a plateau. \gl{I feel we should
  give a conclusion for this paragraph but I am not sure what to say}
\begin{figure}
\begin{center}
\includegraphics[width=0.5\textwidth]{convergence}
\caption{Evolution of the number of hidden units and loss over time on the \texttt{MNIST} dataset for different $\lambda$ values \label{convergence_plot}}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
  \includegraphics[width=0.5\textwidth]{hyper_opt}
\caption{Distributions of the testing accuracy for different training methods, datasets and architectures using random search\label{hyper_opt_res}}
\end{center}
\end{figure}

>>>>

\author{Guillaume Leclerc, Raul Castro Fernandez, Samuel Madden}
\affiliation{%
  \institution{Massachussets Institute of Technology}
}
\email{[leclerc, raulcf, madden]@mit.edu}

%\author{Guillaume Leclerc}
%\authornote{Visiting Student}
%\MITAffiliation
%\email{leclerc@mit.edu}
%
%\author{Raul Castro Fernandez}
%\MITAffiliation
%\email{raulcf@csail.mit.edu}
%
%\author{Samuel Madden}
%\MITAffiliation
%\email{madden@csail.mit.edu}

@@

When designing Neural Networks, finding the appropriate size
%, number of layers
%and number of neuros per layer, 
%(width and depth)
is key. It consists of determining the number of layers and neurons per layer
that lead to the best accuracy.
%In particular, these hyper parameters have a strong correlation with
%over/underfitting \srm{give reference} \gl{I did some litterature review and
%actually this statement is quite controversial, should we rephrase ?}. 
Unfortunately, we have no reliable way to find the best parameters. Decades of
experimentation led to search strategies and heuristics~\cite{Bengio2012a} that prune
the immense space of possible network sizes and suggest range of values to
explore. Examples of search strategies are
%To select the best candidates, researchers have used strategies such
%as 
random search,
%\gl{Should we also move the references here if we remove the
%pargraph from the related work ?}, 
meta-gradient descent~\cite{Pedregosa2016},
gaussian processes~\cite{Bergstra2011a}, and Parzen Estimators~\cite{Bergstra2011a}. 
%to find good parameters without exhaustively
%exploring the entire hyper-parameter space. 
Although these strategies help with
finding a good set of hyper-parameters, they still require a compute-intensive
search of the space.

@@

In this paper we present a method to automatically find an appropriate network
size, thus
%from a single parameter, 
drastically reducing the hyper-parameter optimization time.
%dimensionality. 
The key idea is to
\emph{learn} the right network size at the same time that the network is
learning the main task. For example, for an image classification task, with our
approach we can provide the training data to a network---without sizing it a
priori---and expect to end up with a network that has learnt to classify images
with an accuracy similar to a manually engineered network.
%\gl{too strong, we do not really have
%  any guarantee on the generalization, how about changing it to: to end up with
%  a network that learnt a tradeoff between size and accuracy}. 
This approach has two main benefits. First, we do no longer need to choose a
network size before training. Second, the final network size will be
appropriate for the task at hand, and not larger. 
%Because the sizing occurs
%dynamically, while the
%network is training, we remove the useless neurons 
%implement a new mechanism that removes
%the deactivated neurons to  
%\gl{Should we talk about the
%fact that it automatically gets rid of dead neurons ?}. 
This is important
because oversized networks have a lower inference throughput and higher memory
footprint.

@@

\textbf{Filter Layers: } They have weights in the range $[0,+\infty]$ and are placed after
linear and convolutional layers of traditional deep networks.
%\gl{This is how I
%  used it but with the new implementation we can be more creative and put them
%  anywhere, should we say "usually placed" insdead ?}. 
The \textit{Filter Layer} takes an input of size $\left(B \times C \times D_1
\times \dots \times D_n\right)$, where $B$ is the batch size, $C$ the number of
features (or channels, in the case of convolutional layers), and $D$ any additional dimension. This structure makes
it compatible with fully connected layers with $n=0$ or convolutional layers
with $n=2$. Their crucial property is a parameter $\theta \in \mathbb{R}^C$,
defined as follows:

@@

Where $\circ$ is the 
%Hadamard product (
pointwise multiplication,
%\gl{Should we
%  keep only one of them to be more concise ?}), 
and $\theta$ is expanded in all
dimensions except the second one to match the input size. It is easy to see
that if for any $k$, if $\theta_k \leq 0$, the $k^{\text{th}}$ feature/channel
will forever be $0$. When this happens, we say the Filter layer deactivates the
neuron. These disabled neurons/channels can be removed from the network without
changing its output. Before explaining how that is achieved, we 
% (we describe how we perform this removal below). We 
explain
next how the weights of the Filter Layer are adjusted during training.

@@

%\subsection{The training procedure}
\textbf{Training Procedure: } Once Filter layers are placed in a deep network, 
%To train networks we need start with a substantially oversized network, then we
%insert \textit{Filter Layers}  (usually after every linear or convolutional
%layer except the last one) and we sample their weight from the
%$\text{Uniform}(0, 1)$ distribution. 
we could train it directly and it would be equivalent to a normal neural
network. However, our goal is to find the smallest network with reasonable
performance. We achieve that by introducing sparsity in the parameters of the
\textit{Filter Layers}, thus forcing the deactivation of neurons.
%. Indeed, having a negative component in the $\theta$
%parameter of the filter layer permamently disable its associated feature
%\gl{Maybe redundant ? we talked about that in the previous paragraph} . 
To obtain this sparsity, we simply redefine the loss function:

@@

The additional term $\lambda|\theta|$ introduces sparsity (see Lasso
loss~\cite{Tibshirani1996}).
%This choice comes from the Lasso loss ~\cite{Tibshirani1996} and we use it for
%the same reason: it introduces sparsity. 
The $\lambda$ parameter adjusts how aggressive the network deactivates neurons,
with larger values indicating more aggressive deactivation. 
%The bigger the $\lambda$ the more
%entries in $\theta$ are set to zero, and 
Because the entries in $\theta$ with a value of $0$
correspond to dead neurons, $\lambda$ effectively controls the number of
neurons/channels in the entire network.  
Next, we explain how to implement
ShrinkNets efficiently.

\textbf{Dynamic Network Resizing}
It is possible to reduce the overhead of the training process by removing
neurons as soon as they become deactivated.
%If
%disabled neurons are not quickly removed, the overhead might cause the training
%process to be significantly slower than classic neural networks. 
For that, we implemented a \emph{neural garbage collection} mechanism which
prunes deactivated neurons on-the-fly, therefore reducing the processing time
and memory overhead. To support this feature, it is crucial to understand the
information flow between neurons and layers in the neural network. We achieve
this by representing such information flow as a graph, in which edges indicate
what action each node, i.e., neuron, must take.

\section{Evaluation}

We have implemented ShrinkNets, including the \emph{garbage collection}
mechanism on top of \textit{PyTorch}~\cite{paszke2017automatic}, and we have
made the software open source.\footnote{The code is available as Python/PyTorch
library on \url{http://github.com/mitdbg/fastdeepnets} 
%\gl{Should we rename the repository ?}
}. We have then used our implementation to evaluate two crucial aspects of
ShrinkNets, which we present next.

%This is why we
%want to support what we call garbage collection: removing the weights in the
%layers that are responsible (or using) features that are disabled.  \par To be
%able to catpure this relationship between layers, we augmented \textit{Pytorch}
%~\cite{paszke2017automatic} (the framework we used as the
%foundation of our implementation) with a graph structure very similar to the
%one available in Keras ~\cite{chollet2015keras}.  In this graph, edges are
%effectively event hubs responsible for propagating \textit{feature removal}
%events to its endpoints. ShrinkNet layers are designed to emit and react to
%these events but also propagate the event further if has an impact at the other
%endpoint of the layer (input or output) \gl{What I mean here is: if the event
%  comes from the output, it might be propagated to the inputs, and the other
%  way around}.  \par This event-based implementation coordinated by edges makes
%it very easy to integrate new layers in the library.  We even provide automatic
%wrapping for layers from \textit{PyTorch} if they have no internal state. For
%more complex ones, we provide utilitaries that
%makes the implementation very concise.

%\textbf{Software architecture
%\footnote{The code is available as Python/PyTorch library on \url{http://github.com/mitdbg/fastdeepnets} \gl{Should we rename
%      the repository ?}
%}
%: }\textit{Filter Vectors} can easily be implemented in
%a few lines of code in many existing deep learning frameworks. However, in
%ShrinkNets, we assume that we start with obviously oversized networks. 
%

\subsection{Does ShrinkNet converge?}

%To demonstrate that the approach is viable, we will first show that Shrinking
%Networks converge properly. For this experiment 
We trained a one hidden layer
neural network with one filter layer to control the number of hidden units. We
initialized the models with $10000$ neurons and trained them on \texttt{MNIST}
~\cite{Lecun1998} using different values for $\lambda$.
Figure~\autoref{convergence_plot} shows that
%We
%summarized the results in \autoref{convergence_plot}. We can see on this plot
%is that for a given $\lambda$ the number of hidden units seems to always
%converge to some value. It has two implications: it seems that 
$\lambda$ works as a proxy for the network size,
% as
%we would think, a proxy of the network size 
with bigger $\lambda$ implying smaller
networks. More importantly, the figure helps us confirm that despite the regular
spikes---which are caused by the dynamic disabling of neurons---ShrinkNets
eventually converge.
%we see in the
%loss function that occur when neurons are disabled will eventually disappear
%because the number of neurons reach a plateau. 
%\gl{I feel we should give a
%  conclusion for this paragraph but I am not sure what to say}

\ra{stopped here}

\begin{figure}
\begin{center}
\includegraphics[width=0.9\columnwidth]{convergence}
\caption{Evolution of the number of hidden units and loss over time on the \texttt{MNIST} dataset for different $\lambda$ values \label{convergence_plot}}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
  \includegraphics[width=0.9\columnwidth]{hyper_opt}
\caption{Distributions of the testing accuracy for different training methods, datasets and architectures using random search\label{hyper_opt_res}}
\end{center}
\end{figure}
